### 2014-11-13 locks.tex
Fixed a few typos.
Add facts derived to table.

\begin{toptab}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 199K & 201K & 205K & 196K & 207K \\
SSSP - Coordinated & 964K & 938K & 870K & 886K & 963K \\
SSSP - Buffered & 1M & 1M & 1M & 1M & 1M \\
\hline
HT - Regular & 25M & 32M & 34M & 34M & 36M \\
HT - Coordinated & 51M & 55M & 85M & 80M & 72M \\
HT - Local-Only & 51M & 55M & 79M & 74M & 68M \\
\hline
LBP - Regular & 15M & 17M & 16M & 11M & 8M \\
SBP - Coordinated & 96M & 97M & 87M & 83M & 77M \\
\hline
MiniMax - Regular & 16M & 19M & 19M & 19M & 20M \\
MiniMax - Coordinated & 51M & 51M & 51M & 51M & 52M \\
\hline
N Queens - Regular & 8K & 15K & 16K & 17K & 20K \\
N Queens - Coordinated & 57M & 24M & 675K & 145K & 46K \\
N Queens - Static & 57M & 44M & 1M & 1M & 151K \\
\hline
\end{tabular}
\vspace*{.5ex}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{toptab}

@@

As can be seen in Table~\ref{queue_instructions} all the coordinated
versions of the programs require significantly more queue operations.
The data for this table comes from recording, during execution, the
number of normal queue operations executed (each costs 1), plus the
number of \emph{percolate-up}/\emph{percolate-down} operations
executed for each manipulation of the priority queue.  As expected,
programs which rely on prioritizing the order of execution to obtain
better asymptotic complexity (e.g., SSSP) improve in overall running
time at the cost of more queue operations.  In all cases, coordination
adds a significant number of total queue operations, but the resulting
overhead is more than compensated for by an improved schedule.

@@

We next investigate the effect of coordination on the syncronization
overhead that comes from the use of more locks.  We measured the
\emph{basic locks}, the number of locks acquired
in the original LM virtual machine and \emph{coordination locks}, the
number of locks acquired to perform coordination operations.  We also
measured how often \texttt{lock()} operations that fail to acquired
the lock on the first attempt.  Lock failures happen so infrequently
(less than 1\% of the time) that they cannot be seen on the figures.

>>>>

\begin{dbltab}
   \begin{center}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c | c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} & \multicolumn{5}{c |}{\textbf{Facts Derived}} \\ \cline{2-11}
& 1 & 2 & 4 & 8 & 16 & 1 & 2 & 4 & 8 & 16\\ \hline
\hline
SSSP - Regular & 199.1K & 203.1K & 202.7K & 203.8K & 206.4K & 3.3M & 3.0M & 3.4M & 3.5M & 3.3M \\
SSSP - Coordinated & 959.0K & 924.3K & 891.0K & 966.3K & 1.0M & 2.0M & 2.3M & 2.3M & 2.2M & 2.3M \\
SSSP - Buffered & 1.0M & 1.0M & 1.1M & 1.0M & 1.1M & 2.0M & 2.1M & 2.3M & 2.3M & 2.4M \\
\hline
HT - Regular & 25.6M & 33.3M & 34.8M & 36.1M & 34.8M & 235.7M & 298.3M & 312.1M & 317.5M & 310.3M \\
HT - Coordinated & 51.6M & 55.2M & 85.9M & 79.9M & 75.6M & 129.2M & 183.4M & 254.1M & 242.0M & 252.4M \\
HT - Local-Only & 51.6M & 54.9M & 81.0M & 76.7M & 67.3M & 129.2M & 180.8M & 221.5M & 217.6M & 207.4M \\
\hline
LBP - Regular & 15.7M & 17.8M & 16.0M & 11.3M & 8.5M & 160.2M & 171.0M & 149.9M & 107.5M & 85.1M \\
SBP - Coordinated & 97.2M & 97.4M & 90.0M & 84.3M & 79.2M & 76.8M & 80.7M & 79.2M & 79.0M & 79.3M \\
\hline \hline
MiniMax - Regular & 16.4M & 18.3M & 20.5M & 19.4M & 20.9M & 76.4M & 76.4M & 76.4M & 76.4M & 76.4M \\
MiniMax - Coordinated & 52.0M & 52.0M & 52.0M & 52.0M & 52.0M & 76.4M & 76.4M & 76.4M & 76.4M & 76.4M \\
\hline
N Queens - Regular & 8.4K & 14.4K & 17.7K & 16.6K & 19.3K & 167.1M & 167.1M & 167.1M & 167.1M & 167.1M \\
N Queens - Coordinated & 57.1M & 24.7M & 1.1M & 141.1K & 41.6K & 167.1M & 167.1M & 167.1M & 167.1M & 167.1M \\
N Queens - Static & 57.1M & 47.2M & 3.4M & 1.5M & 209.2K & 167.1M & 167.1M & 167.1M & 167.1M & 167.1M \\
\hline
\end{tabular}
\vspace*{.5ex}
\end{center}
\scap{queue_instructions}{Total number of queue \scare{instructions} and facts derived per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{dbltab}

@@

As can be seen in Table~\ref{queue_instructions} all the coordinated
versions of the programs require significantly more queue operations.
The data for this table comes from recording the number of \emph{queue instructions} and the number of facts derived.
Queue instructions represent the
number of normal queue operations executed (each costs 1) plus the
number of \emph{percolate-up}/\emph{percolate-down} operations
executed for each manipulation of the priority queue.  As expected,
programs which rely on prioritizing the order of execution to obtain
better asymptotic complexity (e.g., SSSP) improve in overall running
time at the cost of more queue operations.  In all cases, coordination
adds a significant number of total queue operations, but the resulting
overhead is more than compensated for by an improved schedule and reduction
in number of facts produced. For MiniMax and N Queens, the number of facts derived
is the same for all configurations, as expected. The performance seen for those programs
arise from reduced memory usage and improved memory locality.

@@

We next investigate the effect of coordination on the synchronization
overhead that comes from the use of more locks.  We measured the
\emph{basic locks}, the number of locks acquired
in the original LM virtual machine and \emph{coordination locks}, the
number of locks acquired to perform coordination operations.  We also
measured how often \texttt{lock()} operations that fail to acquired
the lock on the first attempt.  Lock failures happen so infrequently
(less than 1\% of the time) that they cannot be seen on the figures.

### 2014-11-13 sssp.tex
Add facts derived to table.

Fig.~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
There are some situations where unnecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.
However, we see a reduction of over 40\% in the number of
derived facts when using the coordinated version \textbf{C} over the regular
version \textbf{R}. The results also show that the coordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling.

\begin{topfig}
\vspace*{-5ex}
   \begin{center}
      \subfloat[]{ \begin{tabular}[b]{ | c | c | c | }
         \hline                       
         \textbf{\# T} & \textbf{R} & \textbf{C} \\ \hline \hline
         1 & 333K & 206K \\
         2 & 300K & 210K \\
         4 & 316K & 208K \\
         8 & 328K & 211K \\
         16 & 343K & 212K \\
         \hline  
         \end{tabular}
         \normalsize
      }
      \subfloat[]{\includegraphics[width=5cm]{results/shortest-uspowergrid.png}}
   \end{center}
   \scap{results:sssp_uspowergrid}{Results for SSSP run on the 
   US power grid network.  On the left we show the number of facts
   derived for the regular \textbf{R} and coordinated
   version \textbf{C} using a variable number of threads \textbf{\#
   T}. On the right, we show the scalability of the regular and
   coordinated version. The speedup values are computed using the
   execution time for 1 thread.}
\vspace*{-3ex plus 0pt minus 1ex}
\end{topfig}

>>>>

Fig.~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
The scalability results show that the coordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling. Overall, the coordinated version produces 40\% less facts
derived when compared to the regular version.
Note that there are some situations where unnecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.

\begin{topfig}
\vspace*{-5ex}
   \begin{center}
      \includegraphics[width=5cm]{results/shortest-uspowergrid.png}
   \end{center}
   \scap{results:sssp_uspowergrid}{Scalability results for SSSP run on the 
   US power grid network. The speedup values are computed using the
   execution time for 1 thread.}
\vspace*{-3ex plus 0pt minus 1ex}
\end{topfig}
